Online (Bandit) Policy Selection for EASY-Backfilling

### 2017-03-31 paper/src/paper.tex
paper changes

Providing the computing infrastuctures needed to solve actual complex
problems arising in the various fields of the modern society (including
climate change, health, green energy or security) is a strategic challenge.
The main pillar to address this challenge is to build extreme-scale
parallel and distributed platforms.
The race for always more computing power and storage capacity does not
only lead to sophisticated specific exascale platforms, but the
objective of the community is to design efficient sustained Petascale
platforms.
However, there is still a long way to go and many scientific and
technical problems to solve by the system management software in order
to adapt to such large-scale evolution and the increasing complexity
in both architecture and applications.
The existing job and resource management softwares allow to run tens
thousands jobs on hundreds thousands cores.
They are based on robust policies, however, despite their positive
properties of relative simplicity and their guarantee of no job starvation,
there is still place for improvements.
We propose here to study how to tune the classical EASY-BackFilling
heuristic using two new learning approaches (namely, .

\bigskip
The management generates a huge amount of data.
During the last few years, there was an explosion of the number of works
at the interface of HPC and BigData, dealing with learning algorithms
(including the authors of this work~\cite{learningruntimes}).
Most of these studies target the process of determining more or less
accurately the value of some specific key parameters, with the idea that
learning better estimates should improve the performances of the
resource manager.

\bigskip
Our focus within this work is to learn how the management system behaves
within the usual framework of selecting jobs of the submission queue
under EASY-BF.
Contrarily to most existing studies, which consider only one execution
trace, we extend the learning on a multiple traces basis.
We built a simulator which generates a huge number of partial executions
coming from multiple users collected in various actual traces and HPC
platforms.
The sampling is explained in detail later in the paper.
This was motivated because learning for a specific trace/platform leads
to limited results, customized for one particular instance (then, what
happens if we target another platform?).
The collected data are too small and moreover, this approach is usually
inefficient to deal with extreme unpredictable events.
The sampling generation allow to smooth the data.

%On se place du point de vue de l'admin system...

@@

Our main contribution is to investigate a new way for learning, which
consists in using of a simulator to generate a large data set extracted
from various actual traces.
We propose two variants of the approach: the ideal situation which
relies on the  precise knowledge of data, and the more realistic
assumption of noisy data (which better reflect the uncertainties).
We also propose a new on-line approach based on multi-armed Bandit to
learn very fast how to prioritize the submission and BackFilling queues.
\bigskip

Our main results are:

l'approche basee sur le simulateur marche bien. meme avec des donnes
imprecises.

Le bandit a des resultats raisonnables, il ne necessite que peu de code
supplementaire...

Dans les details :
on montre que une gestion aleatoire des queues est meilleure que FCFS,
et on identifie des politiques qui marchent bien dans plusieurs traces,
etc..

@@

While parallel job scheduling is a old studied theoretical
problem~\cite{leung2004handbook} \textbf{PREFER DROZDOWSKI HERE + SURVEY
FEITELSON-RUDOLPH-SWIEGELSHON}, the practical ramifications, varying
hypotheses, and inherent uncertainty of the problem applied on the HPC
field have driven
practitioners to use \textit{simple} heuristics (and researchers to
study their behavior).
The two most popular heuristics for HPC platforms are EASY~\cite{easy} and
Conservative~\cite{Mu'alem:2001:UPW:380314.380315} backfillings.

@@

\paragraph
\textbf{Data-aware resource management.}

There was a recent focus on leveraging the huge amount of data available in
large scale computing systems in order to improve their behavior. Some works
use collaborative filtering to colocate tasks in clouds by estimating
application interference~\cite{7516031}.  Some others are closer to the
application level and use binary classification to distinguish benign memory
faults from application errors in order to execute recovery algorithms
(see~\cite{fmodeling} for instance).

Several works use \textbf{this method} (LAQUELLE?) in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that
better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.
However, these works do not address the duality between the cumulative and
maximal scheduling costs, as mentionned in~\cite{learningruntimes}.

A recent work~\cite{mlmem} proposes to help
users specify better memory requirements by applying supervised learning
techniques to job metadata.

While these previous works intend to estimate uncertain parameters, they lack
in validating the impact on the system. We consider in this paper a more
pragmatic approach, which is to directly learn a scheduling policy from a given
seach space.

\paragraph
\textbf{Multi-Armed Bandits.}

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards. At every round, an action (represented by
an arm of the Bandit) must be chosen
in a fixed set and the corresponding reward is observed. The goal of a MAB
algorithm is to maximize the total reward obtained in a successive
number of rounds.
There exist a bunch of works that address this problem under a variety of
constraints, The two most popular settings are the original stochastic  case
and the adversarial case. See~\cite{thompson} for the original work on the
stochastic case and~\cite{Auer2002} for the UCB family of algorithms
\textbf{EXPLAIN HERE THE ACRONYM}.
According to our knowledge, the work
of~\cite{Banos} is the earliest known work to us for the adversarial case
see~\cite{nonstoch} for the EXP family of algorithms. We refer to the
review~\cite{bubnow} for a comprehensive overview of the field. While these
algorithms bound the cumulative difference in loss to the best arm (the
regret), they have functional constraints such as the fact that the rewards
should be contained in a range. The heuristic called
\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to
achieve good results in practice, and does not have
this requirement.

@@

This section presents the systems under study and the scheduling problem.  It
first introduces the EASY-Backfilling heuristic and gives
\textcolor{orange}{(fix ugly sentence)} the problem statement.

@@

The problem addressed in this work is the core logic of Resource and Job
Management Systems (RJMS) such as SLURM~\cite{SLURMdocSCHED}, PBS~\cite{PBSdoc},
OAR~\cite{capit2005batch}, Cobalt~\cite{Cobalt}, and more recently Flux~\cite{flux2014}.

@@

In the classical case, these softwares need to execute a set of concurrent
parallel jobs with rigid (known and fixed) resource requirements on a HPC
platform represented by a pool of $m$ identical resources. This is an on-line
problem since the jobs are submitted over time and their characteristics are only known when they are released.
Below is the description and the notations of the characteriscs of job $j$:

\begin{itemize}
  \item Submission date $r_j$ (also called \textit{release date})

@@

The resource requirement $q_j$ of job $j$ is known when the job is submitted at
time $r_j$, while the requested running time $\widetilde{p_j}$ is given by the
user as an estimate. Its actual value $p_j$ is only known \textit{a posteriori}
when the job really completes.  Moreover, the users have incentive to
over-estimate the actual values, since jobs may be ``killed'' if they
surpass the provided value.
\subsection{EASY Backfilling}
\label{sub:easy}

@@

Like other cost metrics, the waiting time is usually considered in its
\textit{cumulative} version, which means that one seeks to minimize the average
waiting time (\textbf{AvgWait}). It is worth noting that the \textbf{MaxWait},
a.k.a the maximal value of the waiting time of all the jobs is also worthy of
interest. Unfortunately, these criteria can sometimes be dual in practice,
making the problem bi-objective. Subsection~\ref{sub:th} will outline our
approach to managing this aspect.

\subsection{Problem Statement}

Our problem statement is: \textbf{How to tune EASY-Backfilling in an online manner?}
\textcolor{orange}{introduce the difference in workloads, the duality of avg and max, and the online problem}

@@

\begin{enumerate}
  \item If the number of allocated or requested cores of a job  exceeds the size of the machine, we remove the job.
  \item If the number of allocated or requested cores is negative, we use the available positive value as the request. If both are negative, we remove the job.
  \item If the runtime or submission time is negative, we remove the job.
\end{enumerate}

@@

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}
  \caption{Share of the policies chosen by Epsilon-Greedy as a function of time.
  The average choice is obtained by resampling the original trace 100 times and
  aggregating by date.}
  \label{fig:mosb}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.6]{figures/mosaic-UniLu-Ga.pdf}
  \caption{Share of the policies chosen by the Noisy policy. The average
  choice is obtained by resampling the original trace 100 times and
  aggregating by date.}
  \label{fig:mosn}
\end{figure}

>>>>

Providing the computing infrastuctures needed to solve actual complex problems
arising in the various fields of the modern society (including climate change,
health, green energy or security) is a strategic challenge.  The main pillar to
address this challenge is to build extreme-scale parallel and distributed
platforms.  The race for always more computing power and storage capacity does
not only lead to sophisticated specific exascale platforms, but the objective
of the community is to design efficient sustained Petascale platforms.
However, there is still a long way to go and many scientific and technical
problems to solve by the system management software in order to adapt to such
large-scale evolution and the increasing complexity in both architecture and
applications.  The existing job and resource management softwares allow to run
tens thousands jobs on hundreds thousands cores.  They are based on robust
policies, however, despite their positive properties of relative simplicity and
their guarantee of no job starvation, there is still place for improvements.
We propose here to study how to tune the ordering of the submission queues in
the frame of the classical EASY-BackFilling family of heuristics using two new
online learning approaches (an approach based on a simulator and an approach
using multi-armed bandit).

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set.  The increasing amount and
diversity of data generated by large scale platforms motivates the community to
study more sophisticated policies that could leverage these data.  During the
last few years, there was an explosion of the number of works at the interface
of HPC and BigData fileds, dealing with learning algorithms .  Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g. job execution
times, memory requirements, etc.) via prediction.  However, only few works
study the impact of such techniques on the performances of the resource
manager.

Contrarily to most existing studies, which consider the learning of specific
parameters of a trace, we provide a complete system that uses feedback from the
system for adapting the whole behavior online.  Our focus within this work is
to learn how the management system behaves within the usual framework of
EASY-BF.  More precisely, we provide strategies for choosing the most
appropriate queue reordering policy used for job selection.  This approach is
assessed by an extensive resampling-based experimental campaign that uses 7
actual traces.

@@

Our main contribution is to investigate two original methods for learning a
good scheduling policy on a per-system basis.  Each method aims to choose
the best queue reordering policy in a fixed search set of semantically diverse
options as fast as possible.

The first method (full feedback) continuously uses simulation on data gathered
from the past and current system workload. In order to reflect the
uncertainties in the data, we also evaluate a "noisy" variant in which
simulations are imprecise.

The second method (bandit feedback) is a simulator-less approach based on a
multi-armed bandit algorithm that derives feedback by measuring system
performance.

Both these methods are evaluated with a specially developped simulator that is
open-sourced and packaged for the community.

\bigskip

Our main results are as follows: The approach based on the simulator provides
very good results, close to the best policy in the search set. In essence, this
means that we are able to recover in an online fashion existing results that
divide average waiting times by factor of \textcolor{orange}{X} to
\textcolor{orange}{X}. Indeed, we show that the required sample sizes for
policy selection in batch schedulers are small enough for workload resampling to
be avoided in practice. The variant based on noisy data, which is closer to
actual conditions of uncertainties on both jobs and platform, behaves also very
well. The alternate approach based on a multi-armed bandit does also provide
reasonable results at a much lower cost, dividing the average waiting times by
a factor of \textcolor{orange}{X} to \textcolor{orange}{X}.

Let us emphasize the main results of this study: While the First-Come,
First-Serve approach to minimizing the waiting times is a common default
strategy, it can absolutely be outperformed.  We show how to automate selection
of a good alternative policy in two cases (depending whether a simulator is
available or not), and provide the difference in performance to be expected
between both situations.

@@

Parallel job scheduling is a old studied theoretical
problem~\cite{Frachtenberg:2009:JSS:1692356} \textcolor{orange}{(cite
feitelson-rudolph-swiegelshon)} whose practical ramifications, varying
hypotheses, and inherent uncertainty of the problem applied on the HPC field
have driven practitioners to use simple heuristics (and researchers to
study their behavior). The two most popular heuristics for HPC platforms are
EASY~\cite{easy} and Conservative~\cite{Mu'alem:2001:UPW:380314.380315}
backfillings.

@@

\subsection{Data-aware resource management.}

There was a recent focus on leveraging the huge amount of data available in
large scale computing platforms in order to improve system performance. Some
works use collaborative filtering to colocate tasks in clouds by estimating
application interference~\cite{7516031}. Some others are closer to the
application level. For instance, ~\cite{fmodeling} uses binary classification
to distinguish benign memory faults from application errors in order to execute
recovery algorithms.

Several works use similar techniques in the context of HPC, in
particular~\cite{Tsafrir_easypp_2005,learningruntimes}, hoping that better job
runtime estimations should improve the scheduling~\cite{chiang_impact_2002}.
Some algorithms estimate runtime distributions model and choose jobs using
probabilistic integration procedures~\cite{Nissimov2008}.  However, these works
do not address the duality between the cumulative and maximal scheduling costs,
as mentioned in~\cite{learningruntimes}. While these previous works intend to
estimate uncertain parameters, we consider here a more pragmatic approach,
which directly learn a good scheduling policy from a given policy space.

Existing work~\cite{jsspp17} takes this approach in an offline manner, by
splitting workload data in a training/testing fashion. This work studied
EASY-Backfilling and shows that the relative performance of some well-known
priority orders for starting jobs differs between workloads, but is relatively
stable troughout time for a given workload.

\subsection{Multi-Armed Bandits.}

A multi-armed bandit (MAB) problem is a sequential allocation problem with
partially observable rewards. At every round, an action (represented by an arm
of the Bandit) must be chosen in a fixed set and the corresponding reward is
observed. The goal of a MAB algorithm is to maximize the total reward obtained
in a successive number of rounds.  There exist a bunch of works that address
this problem under a variety of constraints, The two most popular settings are
the original stochastic  case and the adversarial case. See~\cite{thompson} for
the original work on the stochastic case and~\cite{Auer2002} for the "Upper
Confidence Bound" family of algorithms. According to our knowledge, the work
of~\cite{Banos} is the earliest known work to us for the adversarial case
see~\cite{nonstoch} for the "Exponential-weight" family of algorithms. We refer
to the review~\cite{bubnow} for a comprehensive overview of the field. While
these algorithms bound the cumulative difference in loss to the best arm (the
regret), they have functional constraints such as the fact that the rewards
should be contained in a range. A simple heuristic called
\textit{Epsilon-Greedy} introduced in~\cite{Auer2002} is known to achieve good
results in practice, and does not have this requirement.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%This section presents the systems under study and the scheduling problem.  It
%first introduces the EASY-Backfilling heuristic and gives \todo{fix} the
%problem statement.

@@

@@

In the classical case, the management software needs to execute a set of
concurrent parallel jobs with rigid (known and fixed) resource requirements on
a HPC platform represented by a pool of $m$ identical resources. This is an
on-line problem since the jobs are submitted over time and their
characteristics are only known when they are released.  Below is the brief
description (with the corresponding notations) of the characteristics of job
$j$:

   \item Submission date $r_j$ (also called \textit{release date})

@@

The resource requirement $q_j$ of job $j$ is known when the job is submitted at
time $r_j$, while the requested running time $\widetilde{p_j}$ is given by the
user as an estimate. Its actual value $p_j$ is only known \textit{a posteriori}
when the job really completes.  Moreover, the users have incentive to
over-estimate the actual values, since jobs may be ``killed'' if they surpass
the provided value.

\subsection{Brief description of EASY Backfilling}
\label{sub:easy}

@@

Like other cost metrics, the waiting time is usually considered in its
\textit{cumulative} version, which means that one seeks to minimize the average
waiting time (\textbf{AvgWait}). It is worth noting that the \textbf{MaxWait},
a.k.a the maximal value of the waiting time of all the jobs is also worthy of
interest. Unfortunately, these criteria are antagonistic in nature. Section~\ref{sub:th} will outline our
approach to address the bi-objective aspect of this problem.

@@

\begin{enumerate}

  \item If the number of allocated or requested cores of a job  exceeds the
    size of the machine, we remove the job.

  \item If the number of allocated or requested cores is negative, we use the
    available positive value as the request. If both are negative, we remove
    the job.

  \item If the runtime or submission time is negative, we remove the job.

\end{enumerate}

@@

\begin{figure}[H]
  \centering
  \includegraphics[scale=0.6]{figures/mosaicbandit-UniLu-Ga.pdf}

  \caption{Share of the policies chosen by Epsilon-Greedy as a function of
  time.  The average choice is obtained by resampling the original trace 100
times and aggregating by date.}

  \label{fig:mosb}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[scale=0.6]{figures/mosaic-UniLu-Ga.pdf}

  \caption{Share of the policies chosen by the Noisy policy. The average choice
  is obtained by resampling the original trace 100 times and aggregating by
date.}

  \label{fig:mosn}
\end{figure}
